---
title: "Understanding Gun Violence"
author: "Matthew Perrotta"
date: "11/7/2019"
output: html_document
---

This exercise will serve two purposes: 1) Parse through gun violence data to better understand underlying factors and 2) to build an adequate regression models using common model selection methods.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

### Load Libraries
```{r}
library(tidyverse)
```

# Reading and Cleaning Data

Reading in the .csv file:
```{r}
gv_data = read.csv('./data/gun-violence-data_01-2013_03-2018.csv') %>% 
  janitor::clean_names()

str(gv_data)

length(unique(gv_data$incident_id)) == nrow(gv_data)
```

The dataset `gv_data` consists of `r nrow(gv_data)` observations and `r ncol(gv_data)` columns. Looking at the data, it seems that there a many variables that can be removed right away, such as `source_url`, as they provide no useful information. It should also be noted that many of the variables are categorized as `factor`, when they should either be `character` or `numeric`.

Changing `factor` variables to `character`.
```{r}
fctr.names = gv_data %>% 
  select_if(is.factor) %>% 
  colnames()

gv_data[,fctr.names] = sapply(gv_data[,fctr.names], as.character)

gv_data[sapply(gv_data, function(x) as.character(x)=="")] = NA

str(gv_data)
```

Removing variables and reclassifying some as factors
```{r}
gv_clean = gv_data %>% 
  select(-c(address, incident_url:incident_url_fields_missing, location_description, notes, participant_name, sources)) %>% 
  separate(date, into = c('year', 'month', 'day'), sep = '-') %>% 
  mutate(year = as.factor(year),
         month = as.factor(month),
         day = as.numeric(day),
         n_casualty = n_killed + n_injured)
         
str(gv_clean)

summary(gv_clean)
```


```{r}

```

